---
title: "Stats - Offensive Rating, Defensive Rating, Net Rating"
---

Load libraries
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
```

Read in the data
```{r}
reg_season_stats <- read.csv("data/Prelim2019_RegularSeasonDetailedResults.csv", stringsAsFactors = FALSE)
tourney_stats <- read.csv("data/NCAATourneyDetailedResults.csv", stringsAsFactors = FALSE)
#preview data
head(reg_season_stats)
tail(reg_season_stats)
head(tourney_stats)
tail(tourney_stats)
```

Combine regular season and tournament
```{r}
#distinguish between season and tournament games (no longer needed)
reg_season_stats$post <- 0
tourney_stats$post <- 1
#combine
combined_stats <- bind_rows(reg_season_stats, tourney_stats)
```

Compute tempo per game for winning and losing teams (required for subsequent calculations)
```{r}
combined_stats <- combined_stats %>% mutate(Wposs = WFGA - WOR + WTO + 0.475*WFTA)
combined_stats <- combined_stats %>% mutate(Lposs = LFGA - LOR + LTO + 0.475*LFTA)
head(combined_stats)
tail(combined_stats)
```

Combine winning teams and losing teams stats
```{r}
#duplicate data
combined_statsCompiled <- combined_stats

#change the column names so that they are identical for merging
names(combined_statsCompiled) <- gsub(x = names(combined_statsCompiled), pattern = "WTeam", replacement = "Team")
names(combined_statsCompiled) <- gsub(x = names(combined_statsCompiled), pattern = "LTeam", replacement = "Team")
names(combined_statsCompiled) <- gsub(x = names(combined_statsCompiled), pattern = "W", replacement = "")
names(combined_statsCompiled) <- gsub(x = names(combined_statsCompiled), pattern = "L", replacement = "")

#combine WTeam and LTeam stats
combined_statsCompiled <- bind_rows(combined_statsCompiled[,c(1,3:4,9:21,35,36)],combined_statsCompiled[,c(1,5:6,22:34,35,37)])
head(combined_statsCompiled)
```

Create skeleton to compute Offensive Rating and Defensive Rating
```{r}
#find all unique values by BOTH Season and TeamID
#this will be the skeleton of the final database we want
combined_statsCompiledUnique <- distinct(combined_statsCompiled,Season, TeamID, .keep_all = TRUE)
#select only relevant columns
combined_statsCompiledUnique <- combined_statsCompiledUnique %>% select(Season, TeamID, Score, poss)

#Computing defensive rating requires an additional step:

#def rating as W team
combined_statsCompiledUnique$OppScoreAsW <- NA
combined_statsCompiledUnique$OppPossasW <- NA

#def rating as L team
combined_statsCompiledUnique$OppScoreAsL <- NA
combined_statsCompiledUnique$OppPossasL <- NA

#preview data
head(combined_statsCompiledUnique)
```

Compute score and possession averages (required for Offensive Rating and Defensive Rating)
```{r}
for (i in 1:NROW(combined_statsCompiledUnique)){
    #off rating
    combined_statsCompiledUnique[i,c(3:4)] <- colSums(combined_statsCompiled %>%
      filter(Season == combined_statsCompiledUnique[i,1], TeamID == combined_statsCompiledUnique[i,2]) %>%
      dplyr::select(c(Score, poss)))
    
    #def rating as W team
    combined_statsCompiledUnique[i,c(5:6)] <- colSums(combined_stats %>%
      filter(Season == combined_statsCompiledUnique[i,1], WTeamID == combined_statsCompiledUnique[i,2]) %>%
      dplyr::select(c(LScore, Lposs)))
    
    #def rating as L team
    combined_statsCompiledUnique[i,c(7:8)] <- colSums(combined_stats %>%
      filter(Season == combined_statsCompiledUnique[i,1], LTeamID == combined_statsCompiledUnique[i,2]) %>%
      dplyr::select(c(WScore, Wposs)))
}
```

Compute Offensive Rating, Defensive Rating, and Net Rating
```{r}
#Offensive rating
combined_statsCompiledUnique <- combined_statsCompiledUnique %>% mutate(offRTG = 100* (Score / poss))

#defensive rating
combined_statsCompiledUnique <- combined_statsCompiledUnique %>% mutate(defRTG = 100 * ((OppScoreAsL + OppScoreAsW)/(OppPossasL + OppPossasW)))

#net rating
combined_statsCompiledUnique <- combined_statsCompiledUnique %>% mutate(netRTG = offRTG - defRTG)
```

Select relevant columns
```{r}
combined_statsCompiledUnique <- combined_statsCompiledUnique %>% select(Season, TeamID, offRTG, defRTG, netRTG)
head(combined_statsCompiledUnique)
```

Output to file
```{r}
write.csv(combined_statsCompiledUnique, file="output/stats-rating.csv", row.names = FALSE)
```

