---
title: "Stats - Possession Time Per Game (sec.)"
---

Load libraries
```{r}
library(dplyr)
```

Read in the data
```{r}
Events2010 <- read.csv("data/Events_2010.csv", header = TRUE)
Events2011 <- read.csv("data/Events_2011.csv", header = TRUE)
Events2012 <- read.csv("data/Events_2012.csv", header = TRUE)
Events2013 <- read.csv("data/Events_2013.csv", header = TRUE)
Events2014 <- read.csv("data/Events_2014.csv", header = TRUE)
Events2015 <- read.csv("data/Events_2015.csv", header = TRUE)
Events2016 <- read.csv("data/Events_2016.csv", header = TRUE)
Events2017 <- read.csv("data/Events_2017.csv", header = TRUE)
Events2018 <- read.csv("data/Events_2018.csv", header = TRUE)
```

Combine winning teams and losing teams stats
```{r}
#combine WTeam and LTeam stats
seasonResultsCompiled <- bind_rows(Events2010[,c(2,4:5,8:9)],Events2011[,c(2,4:5,8:9)],Events2012[,c(2,4:5,8:9)],Events2013[,c(2,4:5,8:9)],Events2014[,c(2,4:5,8:9)],Events2015[,c(2,4:5,8:9)],Events2016[,c(2,4:5,8:9)],Events2017[,c(2,4:5,8:9)],Events2018[,c(2,4:5,8:9)])
head(seasonResultsCompiled)
```

Create skeleton for to compute possession
```{r}
#find all unique values by BOTH Season and TeamID
#this will be the skeleton of the final database we want
seasonResultsCompiledUnique <- distinct(seasonResultsCompiled,Season, EventTeamID, .keep_all = TRUE)
```

Compute time between events
```{r}
seasonResultsCompiled$TimeBetweenEvents <- ave(seasonResultsCompiled$ElapsedSeconds, FUN = function(x) c(0,diff(x)))
head(seasonResultsCompiled)

seasonResultsCompiled$TimeBetweenEvents[seasonResultsCompiled$TimeBetweenEvents < 0] <- 0
head(seasonResultsCompiled)

#Create new columns
seasonResultsCompiledUnique$TotalPossessionTime <- NA
```

Compute total possession time per team per season
```{r}
#loop through all the unique values
for (i in 1:NROW(seasonResultsCompiledUnique)){
    seasonResultsCompiledUnique[i,c(6)] <- sum(seasonResultsCompiled %>%
      filter(Season == seasonResultsCompiledUnique[i,1], EventTeamID == seasonResultsCompiledUnique[i,5]) %>%
      dplyr::select(-(Season:EventTeamID)), na.rm = FALSE)
}
```

Read regular season game data
```{r}
WinData <- read.csv("data/RegularSeasonDetailedResults.csv", header = TRUE)
head(WinData)
```

Calculate number of total number of games each team played per season
```{r}
#create required columns
seasonResultsCompiledUnique$Wins <- NA
seasonResultsCompiledUnique$Loses <- NA

for (i in 1:NROW(seasonResultsCompiledUnique)){
    #number of games as winning team
    seasonResultsCompiledUnique[i,c(7)] <- NROW(WinData %>%
      filter(Season == seasonResultsCompiledUnique[i,1], WTeamID == seasonResultsCompiledUnique[i,5]))
    #number of games as losing team
     seasonResultsCompiledUnique[i,c(8)] <- NROW(WinData %>%
      filter(Season == seasonResultsCompiledUnique[i,1], LTeamID == seasonResultsCompiledUnique[i,5]))
}
```

Compute possession time per game
```{r}
seasonResultsCompiledUnique <- seasonResultsCompiledUnique %>%
  mutate(PossessionTimePerGame = TotalPossessionTime / (Wins + Loses))
```

Compute possession time per in seconds
```{r}
seasonResultsCompiledUnique <- seasonResultsCompiledUnique %>%
  mutate(PossessionTimePerGameinMinutes = PossessionTimePerGame / 60)
```

Remove unused columns
```{r}
seasonResultsCompiledUnique$WTeamID <- NULL
seasonResultsCompiledUnique$LTeamID <- NULL
seasonResultsCompiledUnique$ElapsedSeconds <- NULL
seasonResultsCompiledUnique$Wins <- NULL
seasonResultsCompiledUnique$Loses <- NULL
```

Rename columns
```{r}
names(seasonResultsCompiledUnique) <- c("Season", "TeamID", "TotalPossessionTime", "PossessionTimePerGame.S","PossessionTimePerGame.M")
```

Output to file
```{r}
write.csv(seasonResultsCompiledUnique, file = "output/stats-possession.csv", row.names = FALSE)
```


