---
title: "Create Training Data"
---

Load libraries
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
```

### Create training skeleton

Read in the data
```{r}
reg_season_stats <- read.csv("data/Prelim2019_RegularSeasonDetailedResults.csv", stringsAsFactors = FALSE)
tourney_stats <- read.csv("data/NCAATourneyDetailedResults.csv", stringsAsFactors = FALSE)
#preview data
head(reg_season_stats)
tail(reg_season_stats)
head(tourney_stats)
tail(tourney_stats)
```

Combine regular season and tournament
```{r}
#distinguish between season and tournament games
reg_season_stats$post <- 0
tourney_stats$post <- 1
#combine
combined_stats <- bind_rows(reg_season_stats, tourney_stats)
```

Preview the data (tail to ensure that 2019 data merged properly)
```{r}
head(reg_season_stats)
head(tourney_stats)
head(combined_stats)
tail(combined_stats)
```

Create training data skeleton (assign result variable, flip 50% of the data)
```{r}
#Keep season and team columns
training_data <- combined_stats[,c(1,3,5,35)]
#Create result variables
training_data$result <- 1
#rename columns
names(training_data) <- c("Season", "team1_id", "team2_id", "post" ,"result")


#Randomly select 50% of the data
num_samples = dim(training_data)[1]
sampling.rate = 0.5
trainingRows1 <- sample(1:num_samples, sampling.rate * num_samples, replace=FALSE)
training1 <- training_data[trainingRows1, ]
trainingRows2 <- setdiff(1:num_samples,trainingRows1)
training2 <- training_data[trainingRows2, ]
#preview
head(training1)
head(training2)

#flip result for 50% of the data
head(training2)
training2$result <- 0

#re-order and rename
training2 <- training2[,c(1,3,2,4,5)]
names(training2) <- c("Season", "team1_id", "team2_id", "post" ,"result")
head(training2)

#combine regular and flipped data
training_data <- bind_rows(training1, training2)

#remove temp variables
rm(trainingRows1, trainingRows2, training1, training2)
```

### Merge in stats

Read pom
```{r}
medianPOM <- read.csv("output/stats-pom.csv", stringsAsFactors = FALSE)
head(medianPOM)
```

Merge in pom
```{r}

medianPOM1 <- medianPOM
names(medianPOM1)[2] <- "team1_id"
names(medianPOM1)[3] <- paste(sep = "_", "team1", names(medianPOM1)[3])

medianPOM2 <- medianPOM
names(medianPOM2)[2] <- "team2_id"
names(medianPOM2)[3] <- paste(sep = "_", "team2", names(medianPOM2)[3])

# Left outer joins to associate win/loss data
training_data <- merge(x = training_data, y = medianPOM1, by = c("Season", "team1_id"), all.x = TRUE)
training_data <- merge(x = training_data, y = medianPOM2, by = c("Season", "team2_id"), all.x = TRUE)

rm(medianPOM1, medianPOM2)

head(training_data)
```

Read rating
```{r}
netRTG <- read.csv("output/stats-rating.csv", stringsAsFactors = FALSE)
head(netRTG)
```

Merge in Rating
```{r}

netRTG1 <- netRTG
names(netRTG1)[2] <- "team1_id"
names(netRTG1)[3:5] <- paste(sep = "_", "team1", names(netRTG1)[3:5])

netRTG2 <- netRTG
names(netRTG2)[2] <- "team2_id"
names(netRTG2)[3:5] <- paste(sep = "_", "team2", names(netRTG2)[3:5])

# Left outer joins to associate win/loss data
training_data <- merge(x = training_data, y = netRTG1, by = c("Season", "team1_id"), all.x = TRUE)
training_data <- merge(x = training_data, y = netRTG2, by = c("Season", "team2_id"), all.x = TRUE)

rm(netRTG1, netRTG2)

head(training_data)
```

Read tempo
```{r}
tempo <- read.csv("output/stats-tempo.csv", stringsAsFactors = FALSE)
head(tempo)
```

Merge in tempo
```{r}
tempo1 <- tempo
names(tempo1)[2] <- "team1_id"
names(tempo1)[3] <- paste(sep = "_", "team1", names(tempo1)[3])

tempo2 <- tempo
names(tempo2)[2] <- "team2_id"
names(tempo2)[3] <- paste(sep = "_", "team2", names(tempo2)[3])

# Left outer joins to associate win/loss data
training_data <- merge(x = training_data, y = tempo1, by = c("Season", "team1_id"), all.x = TRUE)
training_data <- merge(x = training_data, y = tempo2, by = c("Season", "team2_id"), all.x = TRUE)

rm(tempo1, tempo2)

head(training_data)
tail(training_data)
```

Read possession
```{r}
possession <- read.csv("output/stats-possession.csv", stringsAsFactors = FALSE)

#select relevant columns
possession <- possession %>% dplyr::select(-c(TotalPossessionTime, PossessionTimePerGame.S))
head(possession)
```

Merge in possession
```{r}
possession1 <- possession
names(possession1)[2] <- "team1_id"
names(possession1)[3] <- paste(sep = "_", "team1", names(possession1)[3])

possession2 <- possession
names(possession2)[2] <- "team2_id"
names(possession2)[3] <- paste(sep = "_", "team2", names(possession2)[3])

# Left outer joins to associate win/loss data
training_data <- merge(x = training_data, y = possession1, by = c("Season", "team1_id"), all.x = TRUE)
training_data <- merge(x = training_data, y = possession2, by = c("Season", "team2_id"), all.x = TRUE)

rm(possession1, possession2)

head(training_data)
```

Select final columns

NOTE: Possession was excluded from the final training set so that stage 2 submission would run faster (this choice was made after it was thoroughly tested)
```{r}
training_data <- training_data[,c(1,3,2,4:6,8:10,14,7,11:13,15)]
head(training_data)
```

Output to file
```{r}
write.csv(training_data, file = "output/training-data.csv", row.names = FALSE)
```

