---
title: "Model Selection"
---

Load in libraries
```{r, warning=FALSE, message=FALSE}
library(raster)
library(tidyverse)
```

Read in data
NOTE: possession columns were excluded from training data (see "3-create-training-data.RMD")
```{r}
training_data <- read.csv("output/training-data.csv", stringsAsFactors = FALSE)
head(training_data)
```

Manipulate factorized column
```{r}
training_data$Season <- as.numeric(training_data$Season)
```

Load in library
```{r}
library(SciViews)
```

Various manipulations of variables used in the model selection process. Uncomment as needed.
```{r}
#training_data <- training_data %>% mutate(POMdiff = (team1_medianPOM - team2_medianPOM))
training_data <- training_data %>% mutate(team1_medianPOMajd = 100-4*ln(team1_medianPOM+1)-team1_medianPOM/22)
training_data <- training_data %>% mutate(team2_medianPOMajd = 100-4*ln(team2_medianPOM+1)-team2_medianPOM/22)
training_data <- training_data %>% mutate(POMadjDiff = team1_medianPOMajd - team2_medianPOMajd)
training_data <- training_data %>% mutate(netRTGdiff = (team1_netRTG - team2_netRTG))
training_data <- training_data %>% mutate(offRTGdiff = (team1_offRTG - team2_offRTG))
training_data <- training_data %>% mutate(defRTGdiff = (team1_defRTG - team2_defRTG))
#training_data <- training_data %>% mutate(possdiff = (team1_possessionTimeGameM - team2_possessionTimeGameM))
# training_data <- training_data %>% mutate(team1_pom2 = team1_medianPOM^2)
# training_data <- training_data %>% mutate(team2_pom2 = team2_medianPOM^2)
# training_data <- training_data %>% mutate(team1_offRTG2 = team1_offRTG^2)
# training_data <- training_data %>% mutate(team2_offRTG2 = team2_offRTG^2)
# training_data <- training_data %>% mutate(team1_defRTG2 = team1_defRTG^2)
# training_data <- training_data %>% mutate(team2_defRTG2 = team2_defRTG^2)
# training_data <- training_data %>% mutate(team1_netRTG2 = team1_netRTG^2)
# training_data <- training_data %>% mutate(team2_netRTG2 = team2_netRTG^2)
head(training_data)
```

Train and test up to current year (useful, but was not used moving forward)
```{r}
year <- 2018
year <- year-1
```

Select relevant columns (not needed, see above NOTE)
```{r}
#training_data <- training_data %>% dplyr::select(-c(team1_PossessionTimePerGame.M,team2_PossessionTimePerGame.M))
```

Create logistic regression model and simulate historical competition performance

Bank of variables (copy paste as needed):
-team2_tempo -team1_tempo -team1_offRTG -team2_offRTG -team1_defRTG -team2_defRTG -team1_netRTG -team2_netRTG -team1_netRTG -team2_netRTG -team1_medianPOM -team2_medianPOM -team2_possessionTimeGameM -team1_possessionTimeGameM

```{r}
#create blank variable to store outputs
output=c()

#begin loop of simulating competition results
for (year in 2003:2018){

  #select appropriate training data: (all data up until year - 1) + (season data for year)
  a <- training_data %>% filter(Season %in% c(2003:year))
  b <- training_data %>% filter(Season==year & post==0)
  
  #merge training data
  training <- bind_rows(a,b)
  
  #select appropriate testing data: (tourney data for year)
  testing <- training_data %>% filter(Season==year & post==1)
  
  #remove temporary variables               
  rm(a,b)
  
  #select appropriate columns
  training <- training[,c(5:dim(training)[2])]
  testing <- testing[,c(5:dim(testing)[2])]

  #create logistic regression
  LogisticReg <- glm(result ~ . -team1_medianPOMajd -team2_medianPOMajd -team1_medianPOM -team2_medianPOM  -team1_offRTG -team2_offRTG -team1_defRTG -team2_defRTG -team1_netRTG -team2_netRTG -team2_tempo -team1_tempo, data = training, family = binomial(logit))
  
  #predict
  predictions <-predict(LogisticReg, testing, type = "response")
  
  #clamp
  predictions <- clamp(predictions, lower=0.029, upper=0.971, useValues=TRUE)
  
  #round predictions for misclassification
  roundedPredictions <- round(predictions)
  
  #calculate misclassification
  sizeTest <- dim(testing)[1]
  error = sum(roundedPredictions != testing$result)
  misclassification = error / sizeTest
  
  #calculate logloss evaluation
  library(MLmetrics)
  logloss <- LogLoss(y_pred = predictions, y_true = testing$result)
  
  #print outputs
  print("Year, n, Logloss, Misclassification:")
  print(year)
  print(dim(training[1]))
  print(logloss)
  print(misclassification)
  
  #bind results of current loop to output table
  output <- cbind(output, c(year,dim(training[1]), logloss, misclassification))
}
```

Output to results to file and view details
```{r}
output <- output[-c(3),]
write.csv(output, file = "output/model-selection.csv", row.names = FALSE, col.names = FALSE)

#view logistic regression details
summary(LogisticReg)
print(predictions)
```



